†---
title: "STAT331 Project"
author: "Sijie Jin, Raymond Tan, Yinong Wang"
date: '2018-11-14'
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

# Model Selection

## Pre-fitting data diagnostic

```{r}
# read in data
births.raw = read.csv("chds_births.csv")
```

### Summary of Data

```{r}
head(births.raw, 10)
summary(births.raw)
```

```{r}
# Validating marital
births.raw$marital[!births.raw$marital %in% c(1:5)] = NA
```


```{r}
# Categorize meth
meth.categories = c(rep('Caucasian', 6), 'Mexican', 'African-American', 'Asian', 'Mixed', 'Other')
for (i in 1:length(meth.categories)) {
  births.raw$meth[births.raw$meth == i-1] = meth.categories[i]
  births.raw$feth[births.raw$feth == i-1] = meth.categories[i]
}

# Categorize med
med.categories = c('elementary school', 'middle school', 'high school', 'high school + trade school', 'high school + some college', 'college graduate', 'trade school', 'high school unclear')
for (i in 1:length(med.categories)) {
  births.raw$med[births.raw$med == i-1] = med.categories[i]
  births.raw$fed[births.raw$fed == i-1] = med.categories[i]
}

# Categorize marital
marital.categories = c('married', 'legally separated', 'divorced', 'widowed', 'never married')
for (i in 1:length(marital.categories)) {
  births.raw$marital[births.raw$marital == i] = marital.categories[i]
}

# Categorize income
income.categories = c('under 2500', '2500-4999', '5000-7499', '7500-9999', '10000-12499', '12500-14999', '15000-17499', '17500-19999', '20000-22499', 'over 22500')
for (i in 1:length(income.categories)) {
  births.raw$income[births.raw$income == i-1] = income.categories[i]
}

# Categorize smoke
smoke.categories = c('never', 'smokes now', 'until pregnancy', 'used to, not anymore')
for (i in 1:length(smoke.categories)) {
  births.raw$smoke[births.raw$smoke == i-1] = smoke.categories[i]
}

# Categorize time
time.categories = c('never smoked', 'still smokes', 'during pregnancy', 'less than a year', '1-2 years', '2-3 years', '3-4 years', '5-9 years', 'more than 10 years', 'quit but don’t know when')
for (i in 1:length(time.categories)) {
  births.raw$time[births.raw$time == i-1] = time.categories[i]
}

# Categorize number
number.categories = c('never smoked', '1-4', '5-9', '10-14', '15-19', '20-29', '30=39', '40-60', 'more than 60', 'smoked but don’t know how much')
for (i in 1:length(number.categories)) {
  births.raw$number[births.raw$number == i-1] = number.categories[i]
}

head(births.raw, 10)
```

### NA's and $\pm$Inf's

```{r}
proportion <- sapply(births.raw, function(x) { sum(is.na(x))*100/nrow(births.raw) })
proportion

births.reduced <- births.raw[, !names(births.raw) %in% c('fht', 'fwt')]

births.mice <- mice(births.reduced, method = "cart")
births.clean <- complete(births.mice)
anyNA(births.clean)
```

## Automated Model Selection

### Forward Model Selection

```{r}

```

### Backward Model Selection

```{r}

```

### Stepwise Model Selection

```{r}

```

## Manual Model Selection

### Cross Validation

```{r}

```

### PRESS Statistics

```{r}
#Press Statistic
calPress <- function(M){
  press <- resid(M)/(1-hatvalues(M))
  return(press)
}

press1 <- calPress(M1);
press2 <- calPress(M2);
```



### Akaike Information Criterion
```{r}
# AIC Statistic for M1
AIC1 <- AIC(M1)

#AIC Statistic for M2
AIC2 <- AIC(M2)
```



## Conclusion

# Model Diagnostics

```{r}
#anova(M1, M2)

M0 <- lm(wt ~ 1, data = births.clean)
Mfull <- lm(wt ~ (.)^2, data = births.clean)
Mstart <- lm(wt ~ gestation + meth + mage + mwt + smoke + number, data = births.clean)

# forward
M1 <- step(object = M0,
             scope = list(lower = M0, upper = Mfull),
             direction = "forward",
             trace = FALSE)


# stepward
M2 <- step(object = Mfull, 
              scope = list(lower = M0, upper = Mfull),
              direction = "both", 
              trace = TRUE)

cex <- .8

# studendized residual for M1:



# standardized residual for M1:

res1 <- residuals(M1) 
X1 <- model.matrix(M1) # design matrix
H1 <- X1 %*% solve(crossprod(X1), t(X1)) # Hat matrix
h1 <- diag(H1)
range(h1 - hatvalues(M1))
sigma.hat1 <- sigma(M1)
par(mar = c(4,4,.1,.1))
hist(res1/sigma.hat1, breaks = 50, freq = FALSE, cex.axis = cex,
xlab = "Standardized Residual for M1", main = "")
curve(dnorm(x), col = "red", add = TRUE)

# standardized residual for M2

res2 <- residuals(M2) 
X2 <- model.matrix(M2) # design matrix
H2 <- X2 %*% solve(crossprod(X2), t(X2)) # Hat matrix
h2 <- diag(H2)
range(h2 - hatvalues(M2))
sigma.hat2 <- sigma(M2)
par(mar = c(4,4,.1,.1))
hist(res2/sigma.hat2, breaks = 50, freq = FALSE, cex.axis = cex,
xlab = "Standardized Residual for M2", main = "")
curve(dnorm(x), col = "red", add = TRUE)


```

## Linear Model Assumptions

## Colinearity

## Outliers

# Conclusions

