†---
title: "STAT331 Project"
author: "Sijie Jin, Raymond Tan, Yinong Wang"
date: '2018-11-14'
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

# Model Selection

## Pre-fitting data diagnostic

```{r}
# read in data
births.raw = read.csv("chds_births.csv")
```

### Summary of Data

```{r}
head(births.raw)
summary(births.raw)

proportion <- sapply(births.raw, function(x) { sum(is.na(x))*100/nrow(births.raw) })
proportion

births.reduced <- births.raw[, !names(births.raw) %in% c('fht', 'fwt')]

births.mice <- mice(births.reduced)
births.clean <- complete(births.mice)
```

### NA's and $\pm$Inf's

```{r}

```

## Automated Model Selection

### Forward Model Selection

```{r}

```

### Backward Model Selection

```{r}

```

### Stepwise Model Selection

```{r}

```

## Manual Model Selection

### Cross Validation

```{r}

```

### PRESS Statistics

### Akaike Information Criterion

## Conclusion

# Model Diagnostics

## Linear Model Assumptions

## Colinearity

## Outliers

# Conclusions

