†---
title: "STAT331 Project"
author: "Sijie Jin, Raymond Tan, Yinong Wang"
date: '2018-11-14'
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

# Model Selection

## Pre-fitting data diagnostic

```{r}
# read in data
births = read.csv("chds_births.csv")
```

```{r}
#births
#births[complete.cases(births), ]
```

```{r}
#births1 = births[, !names(births) %in% c('fht', 'fwt', 'income')]
#births1
#births1[complete.cases(births1), ]
```



### Summary of Data

```{r}
head(births)
summary(births)

proportion <- sapply(births, function(x) { sum(is.na(x))*100/nrow(births) })
proportion


exclude <- c("fht","fwt")
include <- setdiff(names(births),exclude)

births <- births[include]


imp.births <- mice(births,method = 'cart' , printFlag=FALSE)

imp.births.complete <- complete(imp.births)







```

### NA's and $\pm$Inf's

```{r}

```

## Automated Model Selection

### Forward Model Selection

```{r}

```

### Backward Model Selection

```{r}

```

### Stepwise Model Selection

```{r}

```

## Manual Model Selection

### Cross Validation

```{r}

```

### PRESS Statistics

### Akaike Information Criterion

## Conclusion

# Model Diagnostics

## Linear Model Assumptions

## Colinearity

## Outliers

# Conclusions

